---
title: "Lab_5_bobby"
author: "Bobby Miyashiro"
date: "10/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Lab 5 objectives

- Learn to parse dates with "liubridate"
- Get counts of observations with count()
- Uncount() rows
- One and two-sample t-test
- Create a heatmap with geom_title()

```{r, include = FALSE}

# Read in data and attach packages
library(tidyverse)
library(here)
library(janitor)

lobster_abundance <- read.csv(here::here("data", "lobster_abundance.csv"),
                              na = "-99999") %>% 
  janitor::clean_names()

```

Use tidyr::uncount() function to convert our lobster data from frequency format to case format

```{r}
lobster_tidy <- lobster_abundance %>% 
  tidyr::uncount(lobster_count)
```

We're only going to consider "site" as our variable of interest

```{r}

ggplot(lobster_tidy, aes(x = site, y = size_mm)) + 
  geom_jitter(aes(color = site),
              alpha = 0.5,
              width = 0.2)

# Histogram
ggplot(lobster_tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill = site)) +
  facet_wrap(~site, scales = "free") # scales equals free allows for scales to be independent for each site

#Quantile Quantile plot
ggplot(lobster_tidy, aes(sample = size_mm)) +
  geom_qq() +
  facet_wrap(~site)
```

#### Convert the 'date' column to class 'Date'

We'll use the 'lubridate' pacakage to convert to Date format, and then to help us easily parse month & year

```{r}

lobster_date <- lobster_tidy %>% 
  mutate(
    date_new = lubridate::mdy(date) #this is convert the dates in the date column to a new class that is unambigous of the date
  )
```

Now let's parse year and month using lubridate::month() and lubridate::year()

```{r}

lobster_parse_date <- lobster_date %>% 
  mutate(
    obs_month = lubridate::month(date_new, label = TRUE), #create a new column of only the month from the date_new column, label = TRUE changes the month number to the month name. Lubridate month knows that the months should be ordered.
    obs_year = lubridate::year(date_new))
    
```

Now let's find counts of observed lobsters based on different hiearchical groupings:

First: let's count lobsters by year and month:

```{r}
lobster_ym <- lobster_parse_date %>% 
  dplyr::count(obs_year, obs_month) #this is to count lobsters by year and month

lobster_y <- lobster_parse_date %>% 
  dplyr::count(obs_year) #this is to only count lobsters by year

lobster_site <- lobster_parse_date %>% 
  dplyr::count(site)

# Having data in tidy format allows us to determine what grouping and counts are important (that's why we use "uncount" earlier)
```

If we want to create a summary table that contains statistics OTHER than counts by group, it's easier to use group_by() + n()

```{r}
lobster_summary <- lobster_parse_date %>% 
  group_by(site) %>% 
  summarize(
    mean_size = mean(size_mm, na.rm = TRUE), # to find mean size of lobsters
    sd_Size = sd(size_mm, na.rm = TRUE), # to find standard deviation of size
    lobster_number = n() #to just find lobster count
  )
```

#### Find confidence intervals

Use t.test() function to find confidence interavls (for one sample) and perform t-tests to compare means of two samples (...this will be covered conceptually in lectures Week 6)

```{r}
ivee_lobster <- lobster_tidy %>% 
  dplyr::filter(site == "IVEE") %>% 
  pull(size_mm)

t.test(ivee_lobster)

```

#### Two-sample t-test to compare means

We want to ask: Is there a significant difference in lobster lengths at Naples and Mohawk reefs?

We've done our necessary exploratory analyses to determine that a 2-sample t-test for means comparison is appropriate.

```{r}

napl_sample <- lobster_tidy %>% 
  filter(site == "NAPL") %>% 
  pull(size_mm) #filter for naples lobsters and pull sizes

mohk_sample <- lobster_tidy %>% 
  filter(site == "MOHK") %>% 
  pull(size_mm)

mn_ttest <- t.test(mohk_sample, napl_sample) # t test are either of these significantly different in either direction.
mn_ttest
```

There is a significant difference in lobster lengths between Naples and Mohawk Reef. (t(df) = statistic, p < 0.001, alpha = 0.05).

```{r}
lobster_mn <- lobster_tidy %>% 
  filter(site %in% c("NAPL","MOHK"))

mn_ttest2 <- t.test(size_mm ~site, data = lobster_mn) #tilde as a function of the different levels of the site column

mn_ttest2 # as you can see the pvalue is small and hte means are quite different, can say that they're two different populations, not just by chance.
```

Here is the p-value: `r mn_ttest2$p.value`

Here is my t-statistic: `r mn_ttest2$statistic`

These back ticks above are inline references that will allow the variables to change as the functions change

#### Now: a heatmap!

```{r}
lobster_ys <- lobster_parse_date %>%
  count(obs_year, site) #count of the observations per year and site

ggplot(data = lobster_ys, aes(x = obs_year, y = site)) +
  geom_tile(aes(fill = n))
  
```

